---
title: "Confirmed COVID-19 Cases Per Country Dashboard"
author: "Marcus McKay Jr. - MPH"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
    ## toc_float: true
  html_notebook: 
    theme: flatly
    toc: true
  html_document:
    theme: flatly
    highlight: tango
    toc: true
runtime: shiny
---

The dashboard visualizes the weekly confirmed COVID-19 cases for each country, organized by year.

**Unit of Counting**: The cases reflected in the graphs is determined by laboratory-confirmed COVID-19 infections reported to the World Health Organization (WHO).

*NOTE: The data reflected in the dashboard may not reflect the true scope of global COVID-19 cases due to inconsistencies in testing strategies, variable case definitions, and under reporting.*

```{r Removing Data From R Environment, message=FALSE, warning=FALSE, include=FALSE}

# Clearing out the R Environment
rm(list = ls())

```

```{r Loading Relevant Packages, message=FALSE, warning=FALSE, include=FALSE}

# Library Relevant Packages
library(tidyverse)
library(haven)
library(dplyr)
library(rlang)
library(ggplot2)
library(shiny)
library(markdown)
library(rmarkdown)
library(jsonlite)
library(bslib)

```

```{r Fetching CSV Data, message=FALSE, warning=FALSE, include=FALSE}

# Fetching the CSV data
WHOnCovid19 <- read.csv("https://ourworldindata.org/grapher/weekly-covid-cases.csv?v=1&csvType=full&useColumnShortNames=true")

# Fetching the metadata
## metadata <- fromJSON("https://ourworldindata.org/grapher/weekly-covid-cases.metadata.json?v=1&csvType=full&useColumnShortNames=true")

```

```{r Cleaning Data and Creating The Plot, message=FALSE, warning=FALSE, include=FALSE}

# Changing the Day variable from character to Date format
WHOnCovid19$date <- as.Date(WHOnCovid19$day, format = "%Y-%m-%d", origin = "1970-01-01")

# Dropping the Day variable
WHOnCovid19 <- WHOnCovid19 %>% select(-c(day))

# Creating the country object using unique values from the Entity variable 
country <- unique(WHOnCovid19$entity)

# Renaming the entity variable
WHOnCovid19 <- WHOnCovid19 %>% rename('country_name' = entity)

# Creating an options widget, using input from the Entity variable and the unique choices from the newly created country variable.
options <- selectInput(
                        'Location'
                       , label = 'Choose A Country'
                       , choices = country
                          )

# Rendering the line plot
renderPlot(
           {
             WHOnCovid19 %>%
                            filter(country_name == country) %>%
                                                               ggplot(aes(x = date, y = weekly_cases)) +
                                                               geom_point(color = 'red') +
                                                               geom_line(color = 'red') +
                                                               labs(##title = 'Weekly Confirmed COVID-19 Cases Per Country', 
                                                                    x = 'Year', y = 'Total Cases') +
                                                               theme_classic()
                      }
                  )

```

```{r echo=FALSE}

# Defining the UI
ui <- fluidPage(
                theme = bs_theme(
                  base_font_size = "0.8rem" # Sets a smaller global font size
                                ),
                # Custom CSS to ensure the app uses the full width of the screen
                tags$head(
                  tags$style(HTML("
                        html,
                        body {overflow-x: hidden;
                                    width: 100%;
                                    margin: 0;
                                    padding: 0;
                                          }
                        .row {margin-left: 0;
                                  margin-right: 0;
                                  }
                        .container-fluid { 
                                      padding: 0 15px; 
                                                  }
                        .citation-text {
                                      font-size: 0.7rem; /* Even smaller for the citation */
                                      color: #555;
                                      margin-top: 20px;
                                              }       
                                  "))
                                ),
                
                                      ##  .container-fluid { 
                                      ##   max-width: 100%; 
                                      ##  margin-left: 0; 
                                      ##   margin-right: 0;
                                      ##   } 
                
                ##titlePanel("Weekly Confirmed COVID-19 Data Dashboard"),
                  sidebarLayout(
                    sidebarPanel(
                                ##  width = 9,
                                ##  height = 9,
                      sliderInput(
                                  inputId = "year_range",
                                  label = "Select Year Range:",
                                  min = 2019,
                                  max = 2025,
                                  value = c(2019, 2025),
                                  sep = ""
                                      ),
                         selectInput("Country", "Select Country:", choices = unique(WHOnCovid19$country_name))
                                          ),
                  mainPanel(
                    ## width = 10,
                    plotOutput("scatter", width = "100%", height = "80vh") 
                    ## ,
                    # The Citation Section
                    ## hr(), # Adds a horizontal line
                    ## div(class = "citation-text",
                    ## "Citation:", tags$br(),
                    ## "World Health Organization (2026) – processed by Our World in Data. ",
                    ## "“Weekly confirmed COVID-19 cases” [dataset]. World Health Organization, ",
                    ## "“COVID-19 Dashboard WHO COVID-19 Dashboard - Daily cases and deaths” [original data]. ",
                    ##   "Retrieved January 8, 2026 from ",
                    ##  tags$a(href = "https://archive.ourworldindata.org/20260109-063243/grapher/weekly-covid-cases.html", 
                    ##         "https://archive.ourworldindata.org/20260109-063243/grapher/weekly-covid-cases.html"),
                    ##  " (archived on January 9, 2026)."
                    )
                  )
                )
              ## )

# Defining the Server
server <- function(input, output) {
                                    output$scatter <- renderPlot(
                                                                  { WHOnCovid19 %>%
                                                                                   filter( country_name == input$Country & 
                                                                                           year(date) >= input$year_range[1] & 
                                                                                           year(date) <= input$year_range[2] ) %>% 
                                                                                                      
                                                                    ggplot(aes(x = date, y = weekly_cases)) +
                                                                    geom_point(color = 'red') +
                                                                    geom_line(color = 'red') +
                                                                    labs(title = paste('COVID-19 Cases for', input$Country)
                                                                         , x = 'Year'
                                                                         , y = 'Weekly Number of Cases') +
                                                                    theme_classic()
                                                                         }
                                                                  , width = "auto", height = "auto")
                                           }

# Running the App
shinyApp(ui = ui, server = server)

```

*Disclaimer*: The data used to construct the line graph was derived from the Coronavirus Source Data webpage of the Our World in Data website. Link to the website is <https://ourworldindata.org/grapher/weekly-covid-cases#reuse-this-work>

Citation: World Health Organization (2026) – processed by Our World in Data. “Weekly confirmed COVID-19 cases” [dataset]. World Health Organization, “COVID-19 Dashboard WHO COVID-19 Dashboard - Daily cases and deaths” [original data]. Retrieved January 8, 2026 from <https://archive.ourworldindata.org/20260109-063243/grapher/weekly-covid-cases.html> (archived on January 9, 2026).

```{r Just Extra Code, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
#I have created a new project directory and used app.R to publish my shiny app to shinyapp.io.

#install.packages("devtools")
#in R console
#library(devtools)

#install.packages("rsconnect")
#in R console
#rsconnect::setAccountInfo(name='mlmj26', token='2393BD692BD5378540F32DE453D7326A', secret='NIgEK1YxEi+gGLdC6ZK68O5d9dbDb93QYzEs+h/C')

#library(rsconnect)

#rsconnect::deployApp('/Users/marcusmckay/Desktop/Grad School TextBooks/Winter 2020/LHS 610/HWs/HW4/directory/assignment_4/assignment_4.Rproj')

# Code Updates:
# 01.04.2026 - a. Updated the URL for live WHO COVID_19 data.  b. Updated code syntax to visualize new COVID-19 variables.
# 03.31.2026 - a. fixed the code to run dashboard in htlm. b. updated the size of graph and text. c.fix margins issue with page and graph, by removing/hiding side panel.

# Defining the UI
## ui <- fluidPage(
                # Making sure the plot and text fit the entire webpage
##                tags$head(
##                  tags$style(
##                    HTML(".main-container {
##                            max-width: 100%;
##                            margin-left: auto;
##                            margin-right: auto;
##                          }")
##                                )
##                                  ),
##              
##              titlePanel("nCOVID-19 Data Dashboard"),
##                sidebarLayout(
##                  sidebarPanel(
##                    sliderInput(
##                                inputId = "year_range",
##                                label = "Select Year Range:",
##                                min = 2019,
##                                max = 2025,
##                                value = c(2019, 2025),
##                                sep = ""
##                                   ),
##                    selectInput("Country", "Select Country:", choices = unique(WHOnCovid19$country_name))
##                                   ),
##                  mainPanel(plotOutput("scatter"))
##                                   )
##                      )

```
